---
title: "Asymmetric tail-dependence in extreme climatic events affects population-extinction risk"
author: "Shyamolina Ghosh, Daniel Reuman"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=1in,right=1in,top=1in,bottom=1in"
output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_maintext.sty
      
tables: True
link-citations: True
urlcolor : blue
indent : True

csl: TheAmericanNaturalist.csl
bibliography: Ref_extrisk.bib
---


```{r setup_MainText, echo=F}
library(rmarkdown)
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
options(scipen = 1, digits = 5) #This option round all numbers appeared in the inline r code upto 5th digit
seed<-101
```

\noindent \emph{Affiliations:}

\noindent Ghosh: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA

\noindent Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA

\newpage

\noindent \textbf{ABSTRACT} 
<!--200 words limit : J Am Nat-->

With global climate change, extreme climatic events (ECEs) become more frequent. These ECEs 
could pose severe impact on the resistance or resilience of a population and/ or community level through
life-history traits of the species. Therefore predicting the impact of extreme events 
on population extinction risk is necessery. We here introduce the potential use of "tail-dependence" 
property of two correlated ecological variables in modeling the extinction risk of 
an age-structured single patch model. We also extend this idea 
to density-dependent metapopulation models. For metapopulation models, 
we show extinction risk not only depends on the nature of tail-dependence of 
spatially correlated noise but is also influenced by the inherent dynamics 
of the population-model. To illaborate, we demonstrate left tail dependent noise across spatially correlated patches 
cause higher extinction risk than the right tail dependent noise, when the model show undercompensatory dynamics.
But in overcompensatory regime, the trend is opposite. Our study paves the path for future research considering this tail-dependence instead of simple correlation in population viability analysis.


\textbf{\textit{Keywords:}} extinction risk, extreme events, tail dependence, density-dependent model

\newpage

<!--***DAN: 
1) Shyamolina,  was reminded during my editing of BIVAN that "tail dependence" has 
a specific mathematical definition (now given in the Background section of BIVAN), and that
is not what we mean here when we say "tail dependence". So we cannot go on calling it "tail
dependence" in this manuscript - that would be mike me calling my house "the white house" - 
there already is one of those, and it is much better known than my house, so we cannot
attempt to hijack the name. So can you change terminology throughout, please? Let's refer to
it using terminology such as "associations in the tails" or 
"tail associations". I would like you to provide a 
definition at an appropriate place early in the manuscript that we will call a pair
of positively associated variables "left-tail associated" if their association in their 
left tails is stronger than their association in their right tails (in a sense to be defined
mathematically in Methods, and then in methods we will define things based on cor_l-cor_u).
We call a pair of positively associated variable "right-tail associated" if their association
in their right tails is stronger than their association in their left tails.

2) You need to start getting this paper ready for submission - if you want to send it to 
the Ecosphere special issue (what do you think of that?) then we have to get moving. Please 
let me know soon, after reading about Ecospere on their website. If you think this is a 
good idea, let me know and I will send a summary of the paper to Jon Walter to make sure he is still willing to have us submit the paper to him, and to make sure the paper is close enough
to the theme of the special issue that he is going for, so that he will be willing to see it
submitted there.

3) Please
add the figure you have mentioned in the text but that does not exist. Please run a spell
check. Please go and check every sentence throughout the manuscript for grammatical 
errors. I think you will find some 
and will be able to correct them if you focus on that - there are many, and by correcting them
you will not only make the paper better you will get better at writing grammatically correct 
sentences in English. 

4) Assuming you want to go for Ecosphere and Jon is OK with us submitting there, please read the "guidelines for authors" section of the Ecosphere website 
and format the paper for Ecosphere. 

5) Please make all the improvements that you can see for yourself
are needed! I suspect if you read the paper carefully now after having worked on something
else for a while you will see many things you can improve!
-->

# Introduction\label{Introduction}

<!--how common are the extreme events?-->
For past few decades, extreme climatic events (ECEs) appeared more common than 
usual due to anthropogenic climate change, e.g. temperature or precipitation extremes,
heat-waves, wild-fire etc. [@ummenhofer2017extreme]. Frequent extreme events 
can shift temperature or precipitation-mean, 
change variability of the distribution, as well as change its 
symmetry or skewness [@hansen2012perception]. Extreme climatic events can pose severe
impacts at the individual, population, community and ecosystem level [@bragazza2008climatic; 
@gutschick2003extreme; @jiguet2011community; @mckechnie2009climate;
@hoover2014resistance]. Therefore, 
understanding the long-term impacts of changes 
in ECEs has been a longstanding interest in ecological research 
[@katz2005statistics; @smith2011ecological; @diez2012will; @moritz2013future; @bailey2016tackling]. 

<!--How extinction risk is related to extreme events : past studies
consequences of extreme events on population life history, community-->
Earlier studies showed ECEs may affect species abundances (species re-ordering) 
<!--***DAN: Shyamolina, I do not know what the parenthesis refers to -->
<!--***DAN: Shyamolina, you have obviously done a lot of reading to situate your manuscript
in the appropriate literature, and that is great!-->
through 
life-history traits, alter community structures through cascading effect on 
higher hierarchical levels and lead to local extinction of 
species [@thomas2004extinction; @edgar2010nino; @moreno2011extreme]. 
<!--Should I mention idea of match-mismatch hypothesis?-->
For example, a study [@chiu2013impact] showed extreme flood affected 
the annal survival of a prey community that in turn affected predators in 
aquatic as well as adjacent riparian habitats via trophic linkages. Another study by 
@frederiksen2008demographic used $43$ years of ring-recovery data 
to examine the demographic impact of ECEs 
through temporal variability in juvenile, immature and adult survival
of European shags on the Isle of May in eastern Scotland.
A relatively recent study [@mcdermott2017sensitivity] investigated the sensitivity 
of $41$ UK butterfly species to four different ECEs (extreme heat, extreme cold, 
extreme precipitation, drought) over $37$ years at the site level, across 
each species' life-stages. Experimental study showed successive ECEs with short time intervals 
could weaken the resistance of plant communities too [@dreesen2014successive]. 
All these studies indicated that environmental stochasticity (through ECEs) 
can lead to demographic stochasticity (through influence on reproductive or 
mortality rate) of different age class of a population making it vulnerable 
to local extinction. 
<!--***DAN: Shyamolina, demogra[hic stochasticity is a specific thing. I suspect you may
just want to say instead something like "All these studies indicated that environmental stochasticity (through ECEs) can make a population vulnerable to extirpation."] -->

<!--Earlier study on metapopulation extinction risk : Large-scale spatial synchrony may also be generated by
other processes such as migration and predation, but regional
stochasticity in the form of spatially correlated weather conditions is
probably the dominant synchronizing mechanism (ref: Nature1998) but what about taildep. effect on extrisk?.-->
In the metapopulation context, regional stochasticity 
can simultaneously affected multiple habitat patches where spatially correlated weather conditions were 
probably the dominant synchronizing mechanism for extinctions [@hanski1998metapopulation]. A 21 years-long
study by @tack2015increasing on the Glanville fritillary butterfly demonstrated that frequency 
of drought during early larval instars increased amplitude and span of spatial synchrony 
over a metapopulation network of 4000 dry meadows and could affect the long-term population viability.
Though earlier literature considered the effect of spatial and/ or temporal correlation in estimating extinction eisk of a metapopulation, but to our knowledge, using the concept of "tail dependence" (i.e. correlation would be stronger during extreme years) to model extinction risk remains unexplored.
<!--***DAN: Shyamolina, you cannot intro tail dependence (please change to calling it
tail association) in the last sentence of a paragraph. The reader does not know what that is, 
and a whole paragraph is needed to intro it.-->

<!--***DAN: Shyamolina, you need to make two major changes in how tail associations
are introdcued. 
  First, do it prior to saying anything about copulas. We have been wanting
to disassociate this paper, to some extent, from BIVAN. Well, probably the best way
to do that is by introducing tail dependence first, before copulas. You can 
describe it intuitively as association in the tails, and you can use pictures to describe it.
You can introduce the partial Spearman correlation (briefly in the Intro, more detail in Methods). 
You will need an explnation of why we 
use Spearman instead of Pearson (talk to me if you do not know, but you may want to read the 
new Background section of BIVAN again first). One does not need to know anything about copulas
to understand tail associations.
  Second, the introduction to copulas that is given is outmoded. See the BAckground section
to BIVAN for improvements. But you do not want to say anything like that much about copulas 
here, you just want to briefly link to the concept, cite some of the Intro work that
has been done (includeing BIVAN and some of the paper BIVAN cites for Intro work),
mention that tail associations are one aspect of copula structure that we have found important,
and introduce any specific things you will need, like the Clayton copula.

Right now this is quite a short paper. You have space to explain things and show pedagogical
figures.
-->

<!--How copula could be relevent in estimating extreme event's 
influence on extinction risk : introduce Taildep. concept-->
We recently [@Ghosh_copula] used this tail dependence feature using "copula" - a statistical 
tool broadly used in modeling dependence between variables - to estimate extinction risk of a density-independent 
metapopulation model. Present manuscript is an extension of this concept 
in estimating the extinction risk beyond the density-independent model 
and even applicable for a simple age-structured population where its vital rates are influenced by ECEs. 
Following we will reintroduce the concept of copula in a manner that follows @Ghosh_copula.

<!--brief intro for copula: I am not sure whether it is needed in this paper or not?-->
<!--***DAN: Shyamolina, to answer your above Q, well, you use Clayton and survival 
Clayton in your simulations, so you probably have to explain copulas enough to explain those.-->
Copula, a rank based measure of association, is much more informative 
than linear correlation measure and a widely applied tool in dependence 
modeling [@joe2014_dependence; @nelsen2006_copula]. Copulas separate the information content of a bivariate dataset,
$(x_n, y_n)$ for $n = 1 \ldots N$, into two non-overlapping parts: the information in the marginal 
distributions (which is not about the association between the variables) and the rest of 
the information (which is solely about the association). The association information is revealed by 
the copula plot ($v_n$ against $u_n$) of the raw variables ($y_n$ vs. $x_n$) in Fig...., where 
$u_n$ is the rank of $x_n$ in the set ${x_1, x_2, \ldots , x_n }$, divided by $N + 1$; 
and $v_n$ is the rank of $y_n$ in the set ${y_1, y_2, \ldots , y_n }$, also divided
by $N + 1$. Here the rank of the smallest element of each sets is considered to be 1 and the rank of
the largest element is $N$. We refer to the $u_n$ and $v_n$ as normalized ranks of the $x_n$ and $y_n$. 
Ranking makes the marginal distributions uniform, isolating only the information on 
association between the variables. 

<!--tail dep. of copula & extinction risk-->
A major benefit of the copula approach is that it can effectively capture the tail dependence. 
Tail dependence, already exemplified in Fig..... (add Fig.), is association between the more extreme 
values of distributions. For two positively associated variables, left-tail 
(respectively, right-tail) dependence is stronger association between
values in the left or lower portions, i. e. for smaller values 
(respectively, the right or upper portions, i.e. for higher values) 
of the two distributions. Stronger dependence in either tail causes asymmetric copula 
structure for the joint distribution of two variables. For example, Clayton and Survival Clayton are 
two copula families 
<!--***DAN: Shyamolina, we cannot just mention these, no one will know what they are.
As I said I think we should intro tail associations without copulas, and then subsequently link 
to copulas afterwards.-->
with asymmetric tail that are used to model exclusively 
left and right tail dependence between variables, respectively.
From ecological perspective, it is reasonable to expect that extreme cold winter during growing season
might reduce the adult reproduction rate as well as larval survival (i.e. stronger 
tail dependence between vital rates when temperature is too cold than normal) and 
thereby inflate the population viability. Similarly, in context of 
density-dependent metapopulation models, if extreme environmental fluctuations across sites 
appear to be as synchronous (left tail dependent or right tail dependent between patches 
with same correlation as in Fig.....) then it could wipe out some metapopulation patches 
during unfavorable years unless they are rescued by the nearby populations. This scenario 
could be more complex and specific of the direction of dependence 
depending on the dynamics of the model (see \nameref{Results}). Though tail-dependence 
property of a copula has been used for risk-analysis in other fields like finance, 
hydrology etc. [@alexander2008market; @li2013; @chen2011flood; @Goswami2018], but to our 
knowledge asymmetric tail dependence have never been considered 
in estimating the effect of ECEs through extinction risk of age structured 
model or density-dependent metapopulations.

<!--our objective ?-->
In this paper, we will explore the potential use of asymmetric tail-dependence 
for understanding the extinction risk of populations towards extreme climatic event
by addressing the following questions: (Q1) Does nature of tail dependence 
(left tail or right tail) in vital rates
of a single-patch age-structured population model affect extinction risk of that population differently?
We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
(Q2) How does environmental noises with asymmetric tail dependence across multiple patches
influence extinction risk of a density-dependent metapopulation model? Our speculation 
is that the general finding from density-independent metapopulation model [@Ghosh_copula]
(higher extinction risk if noises across patches were left-tail dependent than the one 
if noises had right-tail dependence) would hold true only in the under-compensatory regime, but 
should be reverse if the model shows over-compensatory dynamics. 
(Q3) We further hypothesize that when there is no dispersal, the above-said switching should occur
near the critical parameter value where the deterministic model shows transition from 
under-compensatory to over-compensatory dynamics.

# Methods\label{Methods}

<!--stochastic matrix model : age structured-->
To investigate (Q1) from \nameref{Introduction} that how does asymmetric tail 
dependent noise affect the extinction risk of
populations in a single patch age-structured model,
we considered a stochastic age-structured model (Eq. \ref{eq_age_str}) 
with two age classes (egg $E$ and female adult $A$). We assumed
the population was semelperous, i.e., adult ($A$) went through the
reproductive episode only once before death. Eggs and adults in the 
next generation evolved depending on the fecundity of adult ($f_a \geq 0$) 
and survival of eggs ($0 \leq s_e \leq 1$) from present generation.
Fluctuating vital rates ($f_a$ and $s_e$) captured the environmental
stochasticity (Eq. \ref{eq_vitalrates}). We made adult's fecundity 
($f_a$) negative density-dependent with its own population. When $\Omega = \infty$ in 
Eq. \ref{eq_vitalrates}, the projection matrix was independent of populations 
(i.e., Eq. \ref{eq_age_str} was "linear" model then), 
but when $\Omega$ was finite, Eq. \ref{eq_age_str} became "nonlinear".
Fecundity ($f_a$) and survival ($s_e$) rates were random draw from a gamma ($\Gamma(k,\theta)$) and
a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. We generated those distributions 
from iid noises with a varied degree of asymmetric tail-dependence between 
fecundity ($f_a$) and survival ($s_e$) rates. Degree of tail dependence had two broad 
categories: moderate and extreme. Each category had specific direction in tail dependence: 
either in left or in right tail of vital rates. Vital rates ($f_a$ and $s_e$) were
left tail dependent when unusual bad years ("catastrophs") suppressed both vital rates synchronously, 
and right tail dependent when unexpected good years ("bonanzas") boosted both of them. 
 
\begin{equation}
\left({\begin{array}{cc} E \\ A \end{array}}\right)_{t+1} = 
\left(\begin{array}{cc} 0 & f_a\\ s_e & 0 \end{array}\right)
\left(\begin{array}{c} E \\ A \end{array}\right)_{t} \label{eq_age_str}
\end{equation}

\begin{equation}
f_a = \Gamma (k, \theta)\exp(-A/\Omega), \hspace{0.2 cm} s_e = \mathcal{B}(\alpha,\beta) \label{eq_vitalrates}
\end{equation}

We used bivariate Clayton copula to model moderate left-tail and Survival Clayton copula to model 
moderate right tail dependence in the vital rates. Both copula had the same Kendall 
correlation ($=0.6$) but opposite dependence in their respective tails. 
We generated a $N \times 2$ matrix, M, for each of the left (Clayton) and 
right (Survival Clayton) tail-dependent bivariate copulas (see `BiCopSim` function from
`VineCopula` package ), where $N$ is number of years ($=25$) multiplied 
by number of simulations ($=10000$) we used. 
We chose parameters for each copula family using 
`iTau` function from `copula` package so that both copula had
same Kendall's correlation ($\tau = 0.6$). Entries within each clumn of
matrix $M$, then became independent of each others but 
entries between columns had a specific tail dependence
with positive Kendall's correlation ($\tau = 0.6$). 
Similarly, we generated extreme left- (or right-) tail dependent matrix M 
from standard iid (Appendix \ref{SM-extr_tail_dep}) and made the marginals uniform so that
it could mimic an extreme tail dependent copula. 

We made the first column of matrix 
$M$ as gamma ($\Gamma(k,\theta)$) distributed 
with shape parameter $k = 2$ and scale parameter $\theta = 1$.
This distribution represented fecundity of the adult population, $f_a \geq 0$.
For a given $\theta$, increasing $k$ would shift the distribution towards
higher fecundity and reduce the extinction risk if the other parameters
remain unchanged. We took the second column of $M$ and applied a 
beta ($\mathcal{B}(\alpha,\beta)$) transformation. $\alpha =3$ and
$\beta = 2$ were the shape parameters that could control the distribution 
of the survival rates of eggs ($s_e$) ranging in between 0 to 1. 
For each $25$ timesteps from each of $10000$ simulations,
we feed pairwise entries from each row of matrix $M$ as the vital rates 
(fecundity $f_a$ and survival rates $s_e$) 
of the stochastic projrction matrix in model (\ref{eq_age_str}).

<!--calculation of ext risk with age-structured model-->
To calculate extinction risk after each time step, we simulated model (\ref{eq_age_str}) for $25$ years
with $10000$ simulations. For each simulation, we started with initial population $(E,A)_{t=0} = (5,10)$
and following Eq. \ref{eq_age_str} we checked whether after each time step the total population $(E+A)_{t}$
was below a certain extinction threshold (we set it as $2$) or not. If total population was below that threshold then we 
considered population went extinct for that simulation. We also calculated $2.5^{th}$ and $97.5^{th}$ quantiles
of $10000$ simulations with the probability of computed extinction risk at each time step (see [`code`](https://github.com/sghosh89/ERC/blob/master/ext_risk_cop_age_str.R)).

<!--density-dependence metapopulation model-->
We explored (Q2) by examining how nature of the tail dependence of an
environmental variable across space could influence the extinction risk of a 
metapopulation model with density-dependence feedback. We considered six density-dependent 
population-growth models (Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, Austin-Brewer)
in noisy environment (see Table \ref{tab_model_summary}) 
from a metapopulation context.
Models were adapted from earlier work
by @cohen1995unexpected, excluding only two : Varley model (this is not a continuous one) 
and Malthus-Condorcet-Mill model (terms containing logarithm of extinct population does not make sense). 
All models assumed a generalized form : $\vec{P}_{t+1} = D \lambda_t \vec{P_t}$,
where $\vec{P_t}$ was a vector of population densities, 
the $i^{th}$ component $P_t^i$ representing
the density in the $i^{th}$ habitat patch ($i=1,2,\ldots,n$) at time $t$. 
The $n \times n$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
of stochastic density dependence form $\lambda_i(t)$ (see second column 
of Table \ref{tab_model_summary}). For simplicity, 
we considered $n$ patches in each model were arranged, evenly spaced, in a line
with the provision of "local" or "global" dispersal among patches. A n $\times$ n dispersal
matrix, $D$ assumed that a fraction $d$ of 
each population patch dispersed during any given time step, equally distributed to
the two or one nearest neighbors in case of local dispersal. For global dispersal, the fraction
$d$ dispersed equally to other $(n-1)$ patches. We called $d$ as dispersal
rate with $d$ being zero there was no dispersal, and hence no chance to rescue an extinct 
patch through migration from nearby metapopulation patches.

We simulated our metapopulation models with noises $(\epsilon_t^i)$ that were 
independent through time with same spatial correlations between patches (to within sampling error)
for every simulation, but moderately (or extremely) left- and right-tail dependent between patches in
different simulations. For moderately left-tail dependent noise we drew points from a $n$ dimensional 
Clayton copula with same Kendall correlation ($= 0.6$) between any two out of $n$ patches and made 
the marginals normal with zero mean and standard deviation $\sigma$ 
(see the specified value $\sigma$ for each model in 3$^{rd}$ 
column of Table \ref{tab_model_summary}). For moderately 
right-tail dependent noise we flipped ($180^{\circ}$ rotation) the points drawn 
from Clayton copula so that all the statistical properties remained 
same except for the direction of tail-dependence. We generated
extreme left- or right-tail dependent noises via algorithm stated in 
Appendix \ref{SM-extr_tail_dep}. Our models (Table \ref{tab_model_summary}) 
computed extinction risk after $25$ years for $10000$ simulations at $n = 5$ patches
with spatially correlated noises having either moderate or extreme tail 
dependence in left and right-tail.  

<!--say how you set ext. threshold = Pe/10-->
We started with $n$ patches of initial population $P_{t=0}^i = P_e^i, (i = 1,2,\ldots,n)$ where 
$P_e^i$ was the equilibrium value (or similar to carrying capacity concept) 
for each density-dependence model (see 4$^{th}$ column of Table \ref{tab_model_summary}).
We assumed population went extinct after each time step if it was below a 
certain threshold (set to 10\% of initial population $P_e$ for individual model). 

The models stated in Table \ref{tab_model_summary} had one common 
feature: they showed undercompensatory
dynamics (monotonic approach towards stable steady state $P_e$) when
growth parameter $r$ remained below a critical 
level, say, $r_c$. Models showed 
overcompensatory dynamics with oscillatory approach 
towards stable equilibrium $P_e$ when $r$ was beyond $r_c$. 
If $r$ became too high and exceeded $r_{bf}$ - the parameter value responsible for Hopf-bifurcation -
model equilibrium $P_e$ became unstable. Beyond Hopf-bifurcation point, model showed sustained
oscillation and eventually led to chaos with increasing $r$. Appendix \ref{SM-model_stability}
summarized the stability analysis of each model in their deterministic form.

<!--test hypothesis from Q2 : ext risk with density-dependence metapopulation model with r from different regime-->
To test our hypothesis from (Q2) of \nameref{Introduction} for each metapopulation model,
we estimated extinction risk for a wide range of parameter $r$ 
that covered the undercompensatory and overcompensatory regime. We 
chose equispaced values of $r$ from the 95\% quantiles 
of possible range ($r_{min},r_{bf}$), where $r_{min}$ was the minimum value of $r$ 
to get a positive, finite equilibrium $P_e$ (for details, see Appendix 
\ref{SM-model_stability}) to start with ($t=0$). For each chosen $r$,
we calculated extinction risk after $25$ years with each metapopulation model 
where population could disperse locally (or globally) with varying dispersal 
rate $d= 0,0.1,0.2,..0.5$ and extreme (or moderate) left 
tail- and right-tail dependent environmental noise 
in different simulations (for code see [`mydispersal.R`](https://github.com/sghosh89/ERC/blob/master/mydispersal.R)).

<!--test hypothesis from Q3 : switching of ext risk with density-dependence metapopulation model with d=0-->
Given hypothesis from (Q2) was valid - that left tail dependent noise made extinction risk 
higher in undercompensatory regime ($r < r_c$) and
right tail dependent noise made extinction risk higher
in overcompensatory regime ($r > r_c$) - we tested our third hypothesis for (Q3).
We expected the switching to take place near $r_c$, not exactly at $r = r_c$ because the transition point $r_c$
from under to overcompensatory dynamics differed in stochastic environment than its 
deterministic analog depending on the underlying nonlinearity of the model.
We plotted extinction risk along y-axis
for both left tail and right tail dependent noises 
against the varying range of $r$ along x-axis when no dispersal ($d=0$) was considered.
We excluded the possibility of dispersal in this case to avoid any rescue effect among patches. 
For each model with extreme (or moderate) left- and right-tail dependent noises, we calculated 
2.5$^{th}$ and 97.5$^{th}$ quantiles
of 10000 simulations with the probability of computed
extinction risk at each $r$. We compared $r_c$ and the value of $r$ at which 
95\% CI of extinction risk from both noises overlapped
with each other. This could give us an idea for supporting our hypothesis.

# Results\label{Results}

<!--Q1 : age structured results : linear and nonlinear model with asymmetric copula-->

For single patch linear age-structured models, we found higher extinction risk for moderate left-tail 
dependence (modeled by Clayton copula) in fluctuating vital 
rates compared with the one when vital rates were right tail dependent (modeled by Survival Clayton copula).
It was expected because an extreme "bad" year (e.g. with severe cold) can suppress 
the fecundity of adult and survival/ maturation of eggs simultaneously 
and thereby can increase the extinction risk of the population. For non-linear age-structured 
models, we got the similar trend for extinction risks with different tail-dependence (left and right).
Non-linear model showed higher extinction risk than its corresponding 
linear analog as it was evident from Eq. \ref{eq_vitalrates} with finite $\Omega = 100$. 
Fig. \ref{SM-fig_extrisk_agestr_mtd} represents this scenario with two copulas having different 
asymmetric tails but same Kendall's correlation ($\tau = 0.6$). Extinction risks for
two different tail-dependent vital rates were not strinkingly different (e.g. see pair of red and
blue lines in Fig. \ref{SM-fig_extrisk_agestr_mtd}A or B) and we anticipated the difference being partly 
dependent on the degree of tail dependence. Therefore we repeated the same analysis with 
vital rates coming from extreme-tail dependent (left and right in two different simulations)
copulas and got the similar results but with much prominent difference between their extinction risks
(Fig. \ref{fig_extrisk_agestr_etd}). This result answered (Q1) and showed extinction risk of 
a population could be considered as direct manifestation of tail-dependence property 
between vital rates of different age-class
of that population. <!--discussion later: add Tuljapurkar's approximation-->

<!--Q2 : metapopulation results : density dependence with extreme tail dep. copula-->

We found density-dependent metapopulation models in stochastic 
environment showed similar patterns for extinction risk as density-independent 
metapopulation model showed in recent study [@Ghosh_copula]. For each density-dependent 
model listed in Table \ref{tab_model_summary}, we observed higher extinction risk for 
the metapopulations when spatially correlated noises were left-tail dependent than the one with right-tail dependence.
But that observation remained valid when the model parameter - here considered as growth rate $(r)$ - was 
sufficiently low so that under-compensatory dynamics (monotonous approach towards
its stable equilibrium) for the deterministic analog of the model was ensured.
We observed right tail dependent noise caused higher extinction 
risk for the metapopulations than left tail dependent noise when the model 
was in over-compensatory regime (oscillatory approach towards stable equilibrium) for higher 
value of $r$. Figs. \ref{fig_extrisk_allmodel_localdisp_etd} and \ref{SM-fig_extrisk_allmodel_localdisp_mtd} show
variation of extinction risk against local dispersal rate ($d$) for six density dependent metapopulation models 
(see Table \ref{tab_model_summary}) with growth rate $r$ considered either from under-compensatory or 
from over-compensatory regime. We also got similar results with our prediction (see Q2 from \nameref{Introduction})
for global dispersal (Figs. \ref{SM-fig_extrisk_allmodel_globaldisp_etd} and \ref{SM-fig_extrisk_allmodel_globaldisp_mtd}). 
Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1_etd} - \ref{SM-fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}) 
demonstrate how extinction risk changed against local dispersal rate for a whole range
of $r$, whereas Figs. (\ref{SM-fig_ricker_risk_global_disp_scl_1_etd} - 
\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}) 
represent the same but for global dispersal among patches. 
Both case showed the signature of "rescue-effect" (i.e. reduced extinction risk for 
finite dispersal compared with $d=0$).

<!--Q3 : metapopulation results : switching near rc for d=0-->
Following Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1_etd} - 
\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}) we hypothesized that the switching 
between the extinction risk generated by left-tail dependent and right-tail dependent noise 
occurred near the critical parameter $r_c$ (see Q3). For simplicity we only considered 
$d=0$ case, i.e., when there was no dispersal between patches (see \nameref{Methods}). 
Figs. \ref{fig_extrisk_d0_etd} and \ref{SM-fig_extrisk_d0_mtd} demonstrate higher extinction risk 
with left-tail dependent noise for $r<r_c$ and with right-tail dependent noise 
for $r>r_c$. $r_c$ was the vertical line in Figs. \ref{fig_extrisk_d0_etd} and \ref{SM-fig_extrisk_d0_mtd} 
indicating the value of $r$ for each model where their deterministic analog switches 
from under- to over-compensatory dynamics. Though the switching visually appered far away
from the theoretical $r_c$ value for Verhulst and Austin-Brewer model, they were 
consistent with our hypothesis as they had comperatively 
smaller range of $r$ than the other models.


# Discussion\label{Discussion}

<!--summary results:-->
We showed the potential use of "asymmetric" tail dependence in modeling the impact of 
extreme ecological events on population-extinction risk. Our results demonstrated left and right-tail 
dependence in vital rates (or in spatially correlated environmental noises) affected extinction risk differently
in years of extreme events ("catastrophs" and "bonanzas"). Though we could expect left tail dependence 
to show higher extinction risk than the one for right tail dependence with age-structured 
and density dependent metapopulation models, it was not always true. In the over-compensatory regime,
right tail dependent noise caused higher extinction risk than the left tail dependent noise. We showed 
that in absence of dispersal the above-said switching took place near the critical parameter value where 
the deterministic model showed transition from under-compensatory to over-compensatory regime.

<!--Tuljapurkar approximation : linear version of age-structured model in stochastic environments-->
We found left tail dependence in vital rates (induced by ECEs) caused higher extinction risk than 
the right tail dependent ones, but the difference was not remarkable (Fig. \ref{SM-fig_extrisk_agestr_mtd}) 
unless the tail dependence was at its extreme (Fig. \ref{fig_extrisk_agestr_etd}). This finding could be explained 
by the approximation @tuljapurkar1990lecture made for the stochastic log(growth rate) in an iid environment.
The approximation considered growth rate being dependent on the most dominant eigen value of the mean projection matrix 
and the temporal variance of the log(population growth rate) caused by environmental stochasticity - which was based
on the linear covariance measure between any two matrix elements, not on their tail-dependence properties. Therefore,
following @tuljapurkar1990lecture, population growth rate in an iid environment could be approximated
as same for our model (Eq. \ref{eq_age_str}) with left-tail or right-tail dependence.

<!--future study: worthy for PVA, IUCN red list:
1. Identify which life stage or demographic process as management target?
2. Old study (Doak1994, Ferguson1995) showed including correlation btw vital rates for PVA makes a difference, 
   what if we include TD through copula? Multisite age str. model with dispersal, directional change
   in spatio-temporal variation for vital rates.
3. Tdep. with -ve correlation between vital rates?
4. extinction risk how propagated through a food web?
5. Not sure : see McCarthy2000: Most previous models of metapopulation viability have either ignored
correlated extinction of patches, or have included uniform correlations, with the correlation between 
nearby patches being the same as the correlation between distant
patches (Hanski 1994; Tilman et al. 1994; Lindenmayer et al. 1995; McCarthy
1996). 
-->
Although we demonstrated that tail dependence between vital rates of an age structured model can
influence extinction risk, substantial work remains to determine the importance of this effect in the 
following scenarios. First, we demonstrated the effect of tail-dependence on extinction risk 
with a basic age structured (two age stages) model where we considered the total semelperous 
population would go extinct if there would be only one female adult and one egg. But in reality population 
viability analysis (PVA) deals with a little bit higher threshold with several age, stage or size class and a variety 
of populations including semelperous and iteroperous [@morris2002quantitative]. Future studies should include 
tail-dependence effect on extinction risk 
for a multiple age-, size- and stage- structured model and estimate extinction risk 
for each age, size and stage class separately 
so that we can identify (via multisite-PVA) which life-stage or demographic process 
to target for population management. Earlier study by @doak1994modeling showed PVA models with correlation 
between different demographic rates indicated marked difference in predicting confidence interval for 
future population size compared with the model which did not include correlation. But when 
such correlation cannot be estimated directly, more generally where the
forms of statistical dependence among the model's parameters are unknown, making incorrect
assumptions about correlations and dependencies may lead to severe over-estimation 
or under-estimation of extinction risks [@ferson1995correlations]. We believe modeling 
dependencies with copula could help in this case. Second, we did not consider multiple sites of 
age-structured metapopulation which could be linked via dispersal. So, possible future investigation 
would be to explore the viability for the metapopulations if vital rates (e.g. fecundity or 
survival of different stage class) across sites showed specific tail dependence 
with spatially correlated extreme climatic events. Third, we considered tail dependence 
between positively correlated vital rates, but what would be the extinction risk 
and viability of a population if demographic rates were negatively correlated with asymmetric tail dependence. 
We could consider this scenario to model extreme events that kill adults and allow more successful
recruitment of seeds or other propagules. Fourth, following the aforementioned idea, it would be worthy 
to model more complex prey-predator model where asymmetric tail dependent copula could be used to mimic
interspecific-dependence. An earlier study [@carnicer2011widespread] showed extreme events
(i.e., heat waves and droughts) increased mortality and defoliation in southern European forest and its 
dependent herbivorous insects and fungi through trophic linkage. <!--ECEs can even truncate 
multiyear insect outbreak dynamics [@esper20061200].--> We believe copula (may be a multidimensional one) has 
a lot more potential to estimate risk of cascading extinction in model food webs. <!--should I mention about 
match-mismatch hypothesis leads to extinction? our phenology paper in prep.--> Fifth, we considered 
our metapopulation patches were equispaced on a linear chain with equal global dispersal 
rate between patches. But in realistic network models, as inter-patch distances 
can be different, the dispersal rate should depend on distance: dispersal is possible at higher 
rates between nearby patches than the patches which are far away. It is worthy to study 
metapopulation viability with non-uniform spatial correlation (i.e. asymmetric tail dependence) in a 
network with different clusters having various dispersal rates.

 <!--FIG : age structure model results : linear, nonlinear-->
\begin{figure*}[!h]
\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
\caption{Extinction risk against time from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models for $10000$ simulations; shading represents 95\% confidence interval. Fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence. \label{fig_extrisk_agestr_etd}}
\end{center}
\end{figure*}

<!--FIG :  summary : all nonlinear model results for local dispersion : extrisk vs. d-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp_extremenoise.pdf}
\caption{Extinction risks are computed after $25$ years for six nonlinear density-dependence models (Table \ref{tab_model_summary}) with local dispersal among $n=5$ patches and for $10000$ simulations. Extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$. Last two columns represent extinction risk calculated for intrinsic growth rate $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_localdisp_etd}}
\end{center}
\end{figure*}


 <!--FIG : all nonlinear model results with extreme noise, d=0, extrisk vs. r-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_extreme_noise.pdf}\\
\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0_extremenoise.pdf}
\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0_extremenoise.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0_extremenoise.pdf}
\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0_extremenoise.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0_extremenoise.pdf}
\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0_extremenoise.pdf}\\
\caption{Transition of higher to lower extinction risk (calculated after $25$ years) for 
the metapopulations affected by extreme left tail dependent noise across
$5$ multiple patches without dispersal ($d=0$), as intrinsic 
growth rate ($r$) goes from under to over-compensatory regime. Extinction risk for 
extreme right tail dependent noise follows the opposite trend.
Error bars represent the standard errors associated with $10000$
simulations for extinction risk computed at each $r$. Dotted line represents the value of $r$
for each model where their deterministic analog switches from under- to over-compensatory dynamics.
\label{fig_extrisk_d0_etd}}
\end{center}
\end{figure*}

<!--Summary tables for 6 nonlinear models used-->
```{r read_tab_model_summary,echo=F, results='asis',message=F}
library(kableExtra)
library(dplyr)
dt<-readRDS("./Results/tab_model_summary.RDS")
knitr::kable(dt, escape=F,
             format="latex",digits=3, align="c",table.env='table*',
             caption = "Summary of six density dependence metapopulation models considered for each patch; last three columns indicate equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}^{i}$ : transition point from stable to unstable equilibrium of deterministic models), respectively. Superscript $i=1,2,...,n$ indicates index for $i^{th}$ patch.\\label{tab_model_summary}",
             booktabs=T, linesep = "\\addlinespace"
             )%>%column_spec(c(1,3,5,6),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
```


\clearpage

# Literature cited

\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{1pt}
\noindent



