---
title: "Main text : Effect of copula on extinction risk"
author: "Shyamolina Ghosh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_maintext.sty
      
tables: True
link-citations: True
urlcolor : blue
indent : True

csl: TheAmericanNaturalist.csl
bibliography: Ref_extrisk.bib
---


```{r setup_MainText, echo=F}
library(rmarkdown)
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
options(scipen = 1, digits = 5) #This option round all numbers appeared in the inline r code upto 5th digit
seed<-101
```


# Introduction

objectives:

1) Nature of tail dependence in environmental noise (or in vital rates) affects extinction risk of a population differently.
       - Age structure stochastic matrix model : extrisk of left tail dep. > extrisk of right tail dep.
         (This difference is small, could be explained by Tuljapurkar's approx. for linear model, see discussion)
        
       - 6 Density-dependence models with extreme tail dep. in noise (local/ global dispersal)
           - extrisk depends on tail dependnece as well as on model parameters (under-compensatory, over-compensatory)
           - Rescue effect ??
           - H0 : switching occurs near rc (without dispersal)


# Methods

# Results

# Discussion
















 
 <!--FIG : age structure model results : linear, nonlinear-->
\begin{figure*}[!h]
\textbf{ \hspace{5.4 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.9 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_100.pdf}\\
\caption{Extinction risk from single patch age structured $(A)$ linear ($\Omega = \infty$) and $(B)$ nonlinear ($\Omega = 100$) models, fecundity of adults and survival of eggs have either extreme left or extreme right tail dependence, numsims=10000. \label{fig_extrisk_agestr}}
\end{center}
\end{figure*}

<!--FIG :  summary : all nonlinear model results for local dispersion : extrisk vs. d-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp.pdf}
\caption{For each of six nonlinear density-dependence models presented in Table \ref{tab_model_summary} with local dispersal, extinction risk varies as dispersion-rate ($d$) goes from $0$ to $0.5$. Last two columns represent extinction risk calculated for bifurcation parameter $(r)$ taken from under-compensatory and over-compensatory regime of deterministic model analog, respectively.\label{fig_extrisk_allmodel_localdisp}}
\end{center}
\end{figure*}


 <!--FIG : all nonlinear model results with d=0, extrisk vs. r-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend.pdf}\\
\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0.pdf}
\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0.pdf}
\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0.pdf}
\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0.pdf}\\
\caption{Transition of higher to lower extinction risk (calculated after $25$ years) for 
the metapopulations affected by extreme left tail dependent noise across
$5$ multiple patches without dispersal ($d=0$), as intrinsic 
growth rate ($r$) goes from under to over-compensatory regime. Opposite trends
was observed with extreme right tail dependent noise.
Error bars represent the standard errors associated with $10000$
simulations for extinction risk computed at each $r$.
\label{fig_extrisk_d0}}
\end{center}
\end{figure*}

<!--Summary tables for 6 nonlinear models used-->
```{r read_tab_model_summary,echo=F, results='asis',message=F}
library(kableExtra)
library(dplyr)
dt<-readRDS("./Results/tab_model_summary.RDS")
knitr::kable(dt, escape=F,
             format="latex",digits=3, align="c",table.env='table*',
             caption = "Summary of six density dependence models considered for each metapopulation patch with local or global dispersal; last three columns indicated equilibrium population value ($P_e$), critical growth rate ($r_c$) where deterministic analog of models showed transition from under- to over-compensatory dynamics and bifurcation parameter ($r_{bf}$) where $P_e$ became unstable from stable steady state, respectively.\\label{tab_model_summary}",
             booktabs=T, linesep = "\\addlinespace"
             )%>%column_spec(c(1,3,5,6),width="1.4cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.5 cm")
```





